%Untrusted time
%VM exits.
%Cannot move as easily as a VM.
%Performance.
%Monitoring is hard when needed.

Still untrusted parts like the time.
Vm exits are a problem.
Cannot move it as easily as a VM, which is a problem since they target cloud deployment.
The performance penalty is pretty high, it improves security but degrades application performance significantly.
Cannot monitor, when on cloud services this is actually required and very important.\\

What we can take away from this paper is actually that applications deployment changed significantly.
Before, we could deploy on our own trusted servers.
Now multiple mutually untrusted applications are co-located on the host and we have this new bidirectional lack of trust between host and guest.
We need to address this issue.
Relying on hardware is an interesting solution because, compared to software, it is less amenable to bugs and hard to attack or circumvent.
The whole libOS thing is also interesting because it shows that by reimplementing the abstractions that regular applications rely on, we can run unmodified applications.
Also allows the application to verify most of the execution stack (e.g., inspect the libOS).
Although authors claim that this design can be easily ported, I see the lack of support for fork as a problem.
Why untrusted channel when SGX has a mechanism to attest a program?

\adrien{Key insight + how to link to exokernel: separation between physical resources and virtual ones, and separation between policies and mechanisms.}