%Important to see the overhead introduced.
%There is a security incentive, but the performance is also important because it is sacrificed here.
%Running unmodified binaries is really an advantage.
%Disk is still a limitation.
%
% Hard to evaluate on simulator, but looks like main penalties come from X.
% 31 to 54 percent slowdown comparend to VM 
% Enclave crossing is what impacts the most.
% They say it's gonna be acceptable but seems high.
Haven's performance evaluation is done within a simulation that emulates SGX instructions.
EPC access time is artificially modified by reducing the DRAM frequency on the test machine.
Haven's performance is measured against a Microsoft SQL Server benchmark and an Apache HTTP server one.
The measurements are compared with the performance achieved by running the application natively, in a Hyper-V VM, and in Drawbridge.
We note that Haven performs less efficiently than all other configuration for database benchmark, with a 13\% slowdown compared to Drawbridge.
On the HTTP server evaluation, with a dedicated filesystem, Haven is less efficient than other configurations but out-performs Drawbridge thanks to the read intensive nature of the workload that accesses pages located within the buffer cache inside the enclave.
Without the private filesystem, Haven incurs a 40\% slowdown compared to Drawbridge.

The authors experimented with synthetic delays for memory allocation and enclave crossing and reported how these affect Haven's performance.
This method enables to evaluate how these applications react to varying SGX instruction delays.
The database benchmark is unaffected by dynamic memory allocation overheads as a very small number of them occur during the execution.
Enclave crossing overheads are, however, impacting the overall performance.
For the HTTP server benchmark, both overheads seem to have an impact on the performance.

Varying the EPC performance shows that memory-intensive workloads would see their performance decrease, although not proportionally as not all accesses are going to the EPC.
For example, the database benchmark suffered a 21\% slowdown when the memory system was slowed down by a third.

More generally, Haven's authors conservatively expect the performance slow down compared to a virtual machine to range from 31\% to 54\%.

Relying on a recent performance evaluation of SGX, we can deduct more precise costs.