%What is the SGX, what's an enclave, what can you do with it.
%Special pages that cannot be accessed or that trigger notification if accessed from outside of the enclave.
%How much pages can you put in there?\\
%
%Conclusion of this part must be that it's useful but cannot be enough to provide shielded execution.
\adrien{
	That's for the traffic between cache and DRAM.\\

	Cryptographic measurement, sha-256 digest.
	Replay resistent as well!!!

	So the way it works: MEE is part of Memory controller. At boot, it generates random keys, and seizes a region of RAM to store integrity tree. Integrity tree is just a MAC tag of nodes (i.e., pages). Merkel Tree, where only the root is stored in trusted part of the system. Enables to verify everything since has hash of its children.
	Counters for versioning (and avoiding replay attacks.)
}

\adrien{
	Hardware based Attestation.
	Enclave's measurement is a secured record of the content of the enclave's and how it was loaded.
	Attestation of hardware environment and the enclave.
	Enclave has persistent hardware based encryption key.
	user owned special persistent key. Also able to change this.

	Attestation: 1) check code and data, and 2) authority above the enclave.
	1) is basically hash of build log (content, relative position of pages, security flags in pages).
	2) Sealing authority + product ID + version number.

	Signed certificated with expected 1) and public key.
	If everything works, store the public key inside 2).

	Two attestation mechanisms. One local (between two enclaves) and a remote one.
	Local attestation: REPORT (two identities, attributes, hardware TCB, MAC tag).
	MAC is signed with REPORT KEY. Enclave specific. Relies on the MRENCLAVE being known and passed as argument to EREPORT.
	MRENCLAVE along with report will sign with target key.

	Remote needs asymmetric keys.
	Quoting enclave helps with this process.
	Then device specific (private) key signes the MAC -> outputs a Quote.
	Intel Enhanced privacy ID.
	EPID bound to processor firmware (so Quote is signed by the processor).

	Sealing allows to save the enclave to "replay" it later on.

}\\


\adrien{maybe more about attestation and general protection rather than useless details}
The Intel Software Guard Extension (SGX) defines a set of new CPU instructions that allows to create and manipulate \emph{enclaves}.
An enclave is a portion of the user address space protected against unauthorized access or modification by software, even the ones running at a higher privileged level.
Data within an enclave is encrypted and integrity protected.
Any attempt to modify an enclave's data from outside of it results in a fault.

The processor relies on a secure storage, called the \emph{Enclave Page Cache} (EPC), to store and keep track of pages used by an executing enclave and metadata corresponding to these enclaves.
An enclave is created by the \lstinline{ECREATE} instruction, which initializes fields in the corresponding \emph{SGX Enclave Control Structure} (SECS).
SECS are not accessible by software and are stored within the EPC.
Each enclave is associated with a single SECS that holds metadata used by hardware to track the type of a page, the enclave it belongs to, ensure that the page is mapped at the correct virtual address, and accessed with respect to its associated access rights.
Memory pages are added to the enclave using the \lstinline{EADD} instruction.
Thanks to SGX2 extensions, dynamic memory allocation/removal, i.e., adding pages after an enclave's initialization, is rendered possible via a cooperation between the enclave and the host OS.
The \lstinline{EAUG} allows the host to add a new page to the enclave, which in turns needs to acknowledge the new page with \lstinline{EACCEPT} before using it.
In the same way, modifying access permissions and removing a page from an enclave relies on cooperative protocols that involve both the enclave and the host.

SGX threading relies on \emph{Thread Control Structure}s (TCS).
Each enclave can contain one or more TCS's.
A TCS specifies an entry point, \lstinline{OENTRY}, the starting point of the thread, and contains a special protected region, the \emph{State Save Area} (SSA), used to save and restore a thread context upon context switches.
Entering and exiting an enclave is done by respectively calling the \lstinline{EENTER} instruction on a TCS, and the \lstinline{EEXIT} instruction.

Asynchronous exits, on the other hand, are required to allow the underlying operating system to handle exceptions and interrupts.
Upon an asynchronous exit, the current SGX thread state as well as information about the cause of the exit are saved within the associated TCS.
Control can be given back to the enclave by either calling \lstinline{ERESUME}, which will resume the saved context, or \lstinline{EENTER}, which allows the enclave to inspect the exception, on the TCS.
The latter option allows the OS to report exceptions to the enclave.
%Different ways to protect:
%1. data in cache protected by TLB access.
%2. upon modification outside of the enclave, next load triggers a fault.
%3. maintains shadow page table.
%
%ECREATE init a structure in protected memory, EADD allows to add a page to enclave.
%A page allocated by the OS but must come from a specific physical pool (EPC)
%Hardware tracks type of EPC page, enclave it belongs to, checks type, current enclape maps the page at correct va, access okay with access rights.
%Paging is allowed. Prevent rollback attack with versioning number
%
%SGX provides attestation CPU-based. Can be used to prove identity and communicate with other enclaves.
%Processor manufacturer is root of trust for attestation.
%
%TCS to protect the enclave register file.
%EENTER on idle TCS
%Can read write outside of enclave but not execute.
%EEXIT or interrupt or exception. Entry and exit points specified by enclave.
%ERESUME on TCS to get back there. Also, state of enclave is saved and linked to TCS.
%Imperfect because OS can still observe some stuff.
%
%Cooperation between host OS and enclave.
%Allocate: host calls EAUG, enclave must EACCEPT.
%Other instructions to change permissions, require cooperation as well since need TLB shootdowns.
%
%SGX is very low level